### 2018-04-12 ident/docs/draft_2.tex
more figure and results changes

	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations. On applying this method to $v_3$, which is modeled using MWC kinetics, we find that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$. However, when $v_3$ is instead modeled using Convenience kinetics, we find that the parameters are locally identifiable based on the definition we provide in Section \ref{sec:ident}. All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots. Although we could not determine the identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}), $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:parameter_value_v3_var}).

	We also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters. For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Figure \ref{fig:parameter_value_v3_var}a). The same is true of globally structurally identifiable parameters $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is fixed (Figure \ref{fig:parameter_value_v3_var}b). Instead, when Convenience kinetics is used to generate in silico data, under the aforementioned conditions all the parameters are globally structurally and practically identifiable (Figure \ref{fig:parameter_value_v3_var}c and d).

	Similarly, we applied the demonstrated methodology to flux $v_5$ that describes a transcription/translation reaction using Hill kinetics, based on only the available metabolomics and the fluxomics information. When we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system. However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$. Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}. In this instance, the identification required data from two different experiments. Again, just as we see for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable. On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally structurally identifiable parameter. However, only one of the two roots ($K_e^{fdp}(2)$) is practically identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b. Thus, enforcing this criteria results in global practical identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:parameter_value_proof}c).

	\textcolor{red}{summarize and start next section on identifiability here}
	(Figure \ref{fig:ident_v3_var}a),  (Figure \ref{fig:ident_v3_var}b).

	 that require only data from two different experiments: $v_2$ and $v_5$. In applying this method to $v_5$

	In applying this method to $v_5$ we find that the analysis returns two distinct roots for $K_e^{fdp}$, denoted as $K_e^{fdp}(1)$ and $K_e^{fdp}(2)$, thereby making it a locally identifiable parameter.

	Next, we used data from three different experiments to identify the three parameters modeling flux $v_3$. Similar to $v_5$, when the parameters of the Convenience Kinetic rate law modeling $v_3$ are identified, each of the three parameters ($V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$) has two distinct roots that can be written in terms of data from three different experiments. Thus, these parameters are locally identifiable based on the definition we provide in section \ref{sec:ident}.

	First, we would like to point out that

	First, in order to show that the methodology predicts exact values of the parameters that are used to generate the in silico experimental data used for identifiability analysis, the predicted values of parameters modeling fluxes $v_1$ ($k_1^{cat}$ and $K_1^{ac}$), $v_2$ ($V_2^{max}$ and $K_2^{pep}$) and $v_5$ ($V_e^{max}$ and $K_e^{fdp}$) are shown in Figure \ref{fig:parameter_value_proof}. We found that the use of different enzyme kinetic rate law models (Convenience kinetics and Monod-Wyman-Changeaux kinetics) for $v_3$ had no bearing on the ability to estimate the true parameter values when the parameters are globally identifiable.

	For $v_5$, the identifiability

	However, we see from the estimated parameter values for either of the two roots for all three parameters (Figure \ref{fig:parameter_value_v3}a and b) that the total number of distinct values that any of the three parameters can take is less than or equal to the total number of combinations of three experiments from which data is used for identifiability analysis. Also of importance is the fact that the experimental data for this analysis is generated using the MWC model for $v_3$. So, we hypothesized that the uncertainty seen in the parameter estimates from the identifiability analysis could be attributed to the difference between the kinetic model used to generate the data and the model used for identifiability analysis. On using the the Convenience Kinetic model to represent $v_3$ for both data generation as well as identifiability analysis, we see that although the uncertainty in the estimated parameter values is reduced (Figure \ref{fig:parameter_value_v3}c and d), the estimated values are not unique and uncertainty persists. Hence, we conclude that under the current scenario these parameters are structurally locally identifiable and practically non-identifiable.

	We summarize the results on the identifiability of parameters in these and other fluxes in the small network in Figure \ref{fig:ident}, and describe them in the following section.

	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}	
	In Figure \ref{fig:ident}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b), $v_3$ (Figure \ref{fig:ident}c) and $v_5$ (Figure \ref{fig:ident}d) belonging to the network shown in Figure \ref{fig:network}. Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter. We see from Figure \ref{fig:ident} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$) for each of the four fluxes is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law model. We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition. We accordingly summarize that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.

@@

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_v2_v3_v5_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$ when there is no noise in the input experimental data. The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter. The total number of experiments used to identify parameters in $v_1$, $v_2$ and $v_5$ is two. The total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is 240, 421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$. Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters. Results for only one of the two roots indicated by (1) is shown here.}\label{fig:ident}
	\end{figure}

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight, keepaspectratio]{figures/figure1/v1_kcat_v3_2_v5_1_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_3$ and c) $v_5$ \textcolor{red}{(need to change figure)} when there is no noise in the input experimental data. The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter. The total number of experiments used to identify parameters in $v_1$ and $v_5$ is two. The total number of data combinations used to identify parameters in fluxes $v_1$ is 240. 421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$. Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters. Results for only one of the two roots indicated by (2) is shown here. Similarly, results for only one of the two roots of $K_e^{fdp}$, indicated as $K_e^{fdp}(1)$ is shown.}\label{fig:ident_root2}
	\end{figure}

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_1_2_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in the convenience kinetic rate law model of flux $v_3$ when a) $K_3^{pep}$ is held constant, and b) $K_3^{fdp}$ is held constant. Note that the two identifiable parameters in both panel a and panel b are globally identifiable. The degree of identifiability of $V_3^{max}$ and $K_3^{pep}$ are higher when $K_3^{fdp}$ is held constant, than the degree of identifiability of $V_3^{max}$ and $K_3^{fdp}$ when $K_3^{pep}$ is held constant.}\label{fig:ident_v3_var}
	\end{figure}

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_v2_v3_1_v5_2_parameter_value}}
		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$. The parameter values for the only the first root of parameters in $v_3$ ($V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$) and the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) are shown. The variability in the estimated values of $V_1^{max}$ can be explained by the fact that experimental data for identification is generated using the enzyme concentrations as a variable for $v_1$, while the identification of parameters is done by subsuming the enzyme concentration into the enzyme turn over parameter ($k_1^{cat}$) to obtain $V_1^{max}$, and identifying $V_1^{max}$. So, the variability in values of $V_1^{max}$ reflects the variability in the metabolite concentrations (\textit{pep}, \textit{fdp}, \textit{E}) in the experimental data. Similarly, data is generated using the Monod-Wyman-Changeaux model for allosteric regulation for $v_3$, and a Convenience Kinetic rate law model is used for model identification purposes.}\label{fig:parameter_value_v1_vmax}
	\end{figure}

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_k1cat_v3_2_v5_1_parameter_values}}
		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_3$, and c) $v_5$. For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$. The parameter values for the only the second root of parameters in $v_3$ ($V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$) and the first root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(1)$) are shown. Note that $K_e^{fdp}(1)$ is not estimated by any combination of two experiments and $V_e^{max}$ is estimated by all combinations. Data is generated using the Monod-Wyman-Changeaux model for allosteric regulation for $v_3$, and a Convenience Kinetic rate law model is used for model identification purposes.}\label{fig:parameter_value_v1_kcat}
	\end{figure}

>>>>

	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations. On applying this method to $v_3$, which is modeled using MWC kinetics, we find that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$. However, when $v_3$ is instead modeled using Convenience kinetics, we find that the parameters are locally identifiable based on the definition we provide in Section \ref{sec:ident}. All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots. We still could not determine identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}). In the box plot in Figure \ref{fig:parameter_value_v3}a and b we show all the possible estimated parameter values for both roots of all three parameters modeling $v_3$, when experimental data is generated using the MWC model for $v_3$. Due to the numerous estimated parameter values, we can designate these parameters are practically non-identifiable. Although using the Convenience kinetics model for $v_3$ to generate the in silico experimental data leads to a reduction in the uncertainty of the estimated parameters (\ref{fig:parameter_value_v3}c and d), we still end up with non-unique parameter estimates. This leads is to conclude that under the current conditions, flux $v_3$ is structurally locally identifiable and practically non-identifiable.

	However, $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:parameter_value_v3_var}).

	Using Figure \ref{fig:parameter_value_v3_var}, we also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters. For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Figure \ref{fig:parameter_value_v3_var}a). The same is true of globally structurally identifiable parameters $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is fixed (Figure \ref{fig:parameter_value_v3_var}b). Instead, when Convenience kinetics is used to generate in silico data, under the aforementioned conditions all the parameters are globally structurally and practically identifiable (Figure \ref{fig:parameter_value_v3_var}c and d).

	Similarly, we applied the demonstrated methodology to flux $v_5$ that describes a transcription/translation reaction using Hill kinetics, based on only the available metabolomics and the fluxomics information. When we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system. However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$. Thus, similar to $v_3$ earlier, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}. In this instance, the identification required data from two different experiments. Again, just as we see for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable. On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally structurally identifiable parameter. However, only one of the two roots ($K_e^{fdp}(2)$) is practically identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b. Thus, enforcing this criteria results in global practical identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:parameter_value_proof}c).

	\textcolor{red}{summarize and start next section on identifiability here}So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism. Although in some instances (e.g., $v_5$) local identifiability could be resolved to obtain global identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$). Moreover, we also see that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information. We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$. Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to parameter non-identifiability or local identifiability. This can lead to uncertainty in parameter estimates made from in vivo experimental data.

	In the following section, we describe and summarize the results on the identifiability of parameters in these fluxes ($v_1$, $v_2$, $v_3$ and $v_5$) from the network shown in Figure \ref{fig:network}.

	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}	
	In Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a and e), $v_2$ (Figure \ref{fig:ident}b and f), $v_3$ (Figure \ref{fig:ident}c and g) and $v_5$ (Figure \ref{fig:ident}d and h) belonging to the network shown in Figure \ref{fig:network}. Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter. We see from Figure \ref{fig:ident} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$) for each of the four fluxes is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law model. We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition. We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.

@@

	(Figure \ref{fig:ident_v3_var}a),  (Figure \ref{fig:ident_v3_var}b).

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.7\textheight,keepaspectratio]{figures/figure1/v1_V1max_v2_v3_1_v5_mwc_ck_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) and e) $v_1$, b)  and f) $v_2$, c) and g) $v_3$, and d) and h) $v_5$ when there is no noise in the input experimental data. The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, 3360 for $v_3$ and 421 for $v_5$) that can identify each parameter is also specified. $v_1$, $v_2$ and $v_5$ require data from two experiments for analysis, and $v_3$ requires data from three experiments. Results for only one of the two possible roots are shown for $v_3$ and $K_e^{fdp}$ in $v_5$. The results obtained using data from the MWC model for $v_3$ are presented in the panels on the left hand side column and the panels on the right hand column show results obtained using data derived from the Convenience kinetics model for $v_3$.}\label{fig:ident}
	\end{figure}

@@

@@
